---
title: "COVID"
author: "Tryfiates, George"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing Libraries
```{r}
library(readr)
library(summarytools)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(magrittr)
library(janitor)
library(readxl)
```

# Reading the VDH Data
Virginia Dept. of Health
```{r}
vdh <- read_csv("VDH_data.csv")
```

# Cleaning the Data
```{r}
vdh %<>% clean_names()
vdh$report_date <- as.POSIXlt(strptime(vdh$report_date, format = "%m/%d/%Y"))
```

# Summarizing the Deaths by Locality for 2020
Ea
```{r}
vdh %>% filter(report_date == as.POSIXlt(strptime("12/31/2020", format = "%m/%d/%Y"))) %>% group_by(locality) %>% summarise(., "total_covid_deaths" = sum(deaths)) -> vdh2
```
# Reading 2020 Population Data Estimates
UVA Cooper Center's Demographic Research Group releases estimates of Virginia's population broken down by county.
```{r}
pop_data <- read_xls("UVA_pop_data.xls", range = "B8:D140", col_names = FALSE)
pop_overhead <- read_xls("UVA_pop_data.xls", range = "B5:D6")
```

# Cleaning the Data
```{r}
pop_overhead %<>% clean_names()
names(pop_data) <- names(pop_overhead)
```

# Saving the VA total population estimate as its own variable
```{r}
va_total_pop <- pop_overhead$july_1_2020_estimate
# checking that the sum of county populations equal the state total
sum(pop_data$july_1_2020_estimate) == va_total_pop
```
# Sorting the DataFrames by Locality
```{r}
pop_data <- arrange(pop_data, pop_data$locality)
vdh2 <- arrange(vdh2, vdh2$locality)
```

# Setting both to Lowercase
```{r}
vdh2$locality %<>% tolower()
pop_data$locality %<>% tolower()
```

# Checking that the Sort matches
```{r}
data.frame(vdh2$locality, pop_data$locality)
```

# Reordering fairfax and fairfax city in VDH Data
```{r}
vdh2$locality[37]
vdh3 <- vdh2[1:37, ]
vdh3 %<>% add_row(vdh2[39,])
vdh3 %<>% add_row(vdh2[38, ])
vdh3 %<>% add_row(vdh2[40:133,])             
```
# Adding variable for the county's percent of the total statewide COVID-19 deaths
```{r}
vdh3 %<>% mutate("pct_of_total_death" = 100*(vdh3$total_covid_deaths)/sum(vdh3$total_covid_deaths))
```

# Adding a variable for the county's percent of the state's population
```{r}
vdh3 %<>% mutate("pct_of_total_pop" = 100*(pop_data$july_1_2020_estimate/va_total_pop))
```

# Adding a variable for percent change comparing the share of population to the share of COVID-19 deaths
```{r}
vdh3 %<>% mutate("pct_diff_expected_dist" = 100*((pct_of_total_death - pct_of_total_pop)/pct_of_total_pop))
```

# Viewing the Dataframe
```{r}
vdh3
```

# Largest Positive % Change
```{r}
vdh3 %>% filter(pct_diff_expected_dist > 200 )
```
# Largest Negative % Change 
```{r}
vdh3 %>% filter(pct_diff_expected_dist < -67)
```
# Summary of Data
```{r}
view(dfSummary(vdh3))
summary(vdh3$pct_diff_expected_dist)
```

# Importing longitude and latitude data for Virginia counties
```{r}
va_map <- map_data("county", "virginia")
```

# Matching DataFrames
```{r}
s_region<- unique(va_map$subregion)
setdiff(vdh3$locality, s_region)
```

```{r}
setdiff(s_region, vdh3$locality)
```
Yep, these are franklin, richmond, and roanoke *county*. 
```{r}
va_map %>% filter(subregion == c("franklin", "richmond", "roanoke")) -> charl
ggplot(data = va_map)+
  geom_polygon(aes(x=long, y=lat, group = group), color = "black", fill="grey") +
  geom_polygon(data = charl, aes(x=long, y= lat, group=group), fill="pink")+
  coord_fixed(1.3)+
  guides(fill=FALSE)+
  theme_nothing()
```
# Editing county names to match
```{r}
vdh3$locality[vdh3$locality=="richmond county"] <- "richmond"
vdh3$locality[vdh3$locality == "franklin county"] <- "franklin"
vdh3$locality[vdh3$locality == "roanoke county"] <- "roanoke"
```

# Merging the DataFrames
```{r}
va_county_fill <- left_join(va_map, vdh3, by=c("subregion"="locality"))
```

# Importing Colorblind Palette
```{r}
library(viridis)
```
# Saving a Dataframe of longitude and latitude for the state outline
```{r}
va_state <- map_data("state", "virginia")
```

# Visualizing the Final Data with My Variable
```{r}
ggplot(data = va_county_fill)+
  labs(title = "Virginia", subtitle = "Comparing County-wise Share of COVID-19 Deaths against Share of VA Population", caption = "Data Sources: VDH + UVA Weldon Cooper Center")+
  geom_polygon(aes(x=long, y=lat, group = group, fill=pct_diff_expected_dist), 
               color = "white") +
  scale_fill_viridis_b(option = "A", direction = -1, name="Percent Change")+
  geom_polygon(data = va_state, aes(x=long, y=lat, group=group), 
               fill=NA, color=viridis(1, begin = 0.2, alpha = 0.2))+
  coord_fixed(1.3)+
  theme_fivethirtyeight()+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust=0.5, face = "italic"), 
        legend.position = c(0.08,0.62), legend.direction = "vertical",
        legend.title = element_text(size = 10.5, face = "bold"), 
        legend.text = element_text(size = 10, face = "italic"),
        text = element_text(family = "sans", size = 12), panel.grid = element_blank(),
        axis.title = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), line=element_blank(), 
        panel.grid.major = element_blank(), plot.caption = element_text(size=7)
        )
```
Basically, if a county has a percent change of 100%, that means that its share of COVID-19 deaths is double its share of the Virginia population.

# Saving the Plot
```{r}
ggsave("VA_countywise_mortality_share_i_excess.png", dpi = 1068)
```
# Playing with More Colors and Alternative Text
```{r}
ggplot(data = va_county_fill)+
  labs(
    title = "Pandemic Proportionality, Virginia 2020", 
    subtitle = 
         "Were the county shares of statewide COVID-19 deaths proportional to its share of population?", 
    caption = "Data Sources: VDH + UVA Weldon Cooper Center"
    )+
  geom_polygon(
    aes(x=long, y=lat, group = group, fill=pct_diff_expected_dist), 
    color = "white"
    )+
  scale_fill_viridis_b(
    
    direction = 1, 
    name="Percent Change"
    )+
  geom_polygon(
    data = va_state, 
    aes(x=long, y=lat, group=group),
    fill=NA, 
    color=viridis(1, begin = 0.2, alpha = 0.2)
    )+
  coord_fixed(1.3)+
  theme_fivethirtyeight()+
  theme(
    plot.title = element_text(hjust = 0.5), 
    plot.subtitle = element_text(hjust=0.5, face="italic"), 
    legend.position = c(0.08,0.62), legend.direction = "vertical",
    legend.title = element_text(size = 10.5, face = "bold"), 
    legend.text = element_text(size = 10, face = "italic"),
    text = element_text(family = "sans", size = 12), 
    panel.grid = element_blank(),
    axis.title = element_blank(), 
    axis.text = element_blank(),
    axis.ticks = element_blank(), 
    line=element_blank(), 
    panel.grid.major = element_blank(), 
    plot.caption = element_text(size=7)
    )
```

```{r}
ggsave("VA_countywise_mortality_share_x_blue_alttext.png", dpi = 1068, width = 8)
```

```{r}
ggplot(data = va_county_fill)+
  labs(
    title = "Pandemic Proportionality, Virginia 2020", 
    subtitle = 
         "Were the county shares of statewide COVID-19 deaths proportional to its share of population?", 
    caption = "Data Sources: VDH + UVA Weldon Cooper Center"
    )+
  geom_polygon(
    aes(x=long, y=lat, group = group, fill=pct_diff_expected_dist), 
    color = "white"
    )+
  scale_fill_viridis_b(
    option = "A",
    direction = 1, 
    name="Percent Change"
    )+
  geom_polygon(
    data = va_state, 
    aes(x=long, y=lat, group=group),
    fill=NA, 
    color=viridis(1, begin = 0.2, alpha = 0.2)
    )+
  coord_fixed(1.3)+
  theme_fivethirtyeight()+
  theme(
    plot.title = element_text(hjust = 0.5), 
    plot.subtitle = element_text(hjust=0.5, face="italic"), 
    legend.position = c(0.08,0.62), legend.direction = "vertical",
    legend.title = element_text(size = 10.5, face = "bold"), 
    legend.text = element_text(size = 10, face = "italic"),
    text = element_text(family = "sans", size = 12), 
    panel.grid = element_blank(),
    axis.title = element_blank(), 
    axis.text = element_blank(),
    axis.ticks = element_blank(), 
    line=element_blank(), 
    panel.grid.major = element_blank(), 
    plot.caption = element_text(size=7)
    )
```

```{r}
ggsave("VA_countywise_mortality_share_x_magma_alttext.png", dpi = 1068, width = 8)
```
```{r}
ggplot(data = va_county_fill)+
  labs(
    title = "Pandemic Proportionality, Virginia 2020", 
    subtitle = 
         "Were the county shares of statewide COVID-19 deaths proportional to its share of population?", 
    caption = "Data Sources: VDH + UVA Weldon Cooper Center"
    )+
  geom_polygon(
    aes(x=long, y=lat, group = group, fill=pct_diff_expected_dist), 
    color = "white"
    )+
  scale_fill_viridis_b(
    option = "A",
    direction = -1, 
    name="Percent Change"
    )+
  geom_polygon(
    data = va_state, 
    aes(x=long, y=lat, group=group),
    fill=NA, 
    color=viridis(1, begin = 0.2, alpha = 0.2)
    )+
  coord_fixed(1.3)+
  theme_fivethirtyeight()+
  theme(
    plot.title = element_text(hjust = 0.5), 
    plot.subtitle = element_text(hjust=0.5, face="italic"), 
    legend.position = c(0.08,0.62), legend.direction = "vertical",
    legend.title = element_text(size = 10.5, face = "bold"), 
    legend.text = element_text(size = 10, face = "italic"),
    text = element_text(family = "sans", size = 12), 
    panel.grid = element_blank(),
    axis.title = element_blank(), 
    axis.text = element_blank(),
    axis.ticks = element_blank(), 
    line=element_blank(), 
    panel.grid.major = element_blank(), 
    plot.caption = element_text(size=7)
    )
```

```{r}
ggsave("VA_countywise_mortality_share_x_magmalite_alttext.png", dpi = 1068, width = 8)
```

```{r}
ggplot(data = va_county_fill)+
  labs(
    title = "Pandemic Proportionality, Virginia 2020", 
    subtitle = 
         "Were the county shares of statewide COVID-19 deaths proportional to its share of population?", 
    caption = "Data Sources: VDH + UVA Weldon Cooper Center"
    )+
  geom_polygon(
    aes(x=long, y=lat, group = group, fill=pct_diff_expected_dist), 
    color = "white"
    )+
  scale_fill_viridis_b(
    option = "A",
    direction = -1, 
    name="Percent Change"
    )+
  geom_polygon(
    data = va_state, 
    aes(x=long, y=lat, group=group),
    fill=NA, 
    color=viridis(1, begin = 0.2, alpha = 0.2)
    )+
  coord_fixed(1.3)+
  theme_fivethirtyeight()+
  theme(
    plot.title = element_text(hjust = 0.5), 
    plot.subtitle = element_text(hjust=0.5), 
    legend.position = c(0.08,0.62), legend.direction = "vertical",
    legend.title = element_text(size = 10.5, face = "bold"), 
    legend.text = element_text(size = 10, face = "italic"),
    text = element_text(family = "sans", size = 12), 
    panel.grid = element_blank(),
    axis.title = element_blank(), 
    axis.text = element_blank(),
    axis.ticks = element_blank(), 
    line=element_blank(), 
    panel.grid.major = element_blank(), 
    plot.caption = element_text(size=7)
    )
```

```{r}
ggsave("VA_countywise_mortality_share_magmalite_alttext.png", dpi = 1068, width = 8)
```

